Isomorphisms as Groupoids: \\ Homotopy Type Theory and Fractional Types

### 2013-11-03 f.tex
talk
f

@@

>>>>

A permutation is represented as

\begin{verbatim}
(1 2 3 4 5)
(2 5 4 3 1)
\end{verbatim}

We can decompose it in:

\begin{verbatim}
2

and

1 with paths (2 3 4 5)
             (5 4 3 1)
\end{verbatim}

This is the decomposition of the function \verb|A -> A| into
\verb|A * 1/A|. A problem still with cardinality: there is exactly one
element into the latter type but presumably several functions in the
former. Perhaps all these functions are equivalent somewho up to homotopy?

@@

{--

-- In our perspective isomorphisms are represented by paths; there are
-- 1! isomorphims from ? to itself so we expect exactly one path in
-- the space ? o-> ?, i.e., one path from ? to itself, i.e., the space
-- (? ? ?) should be isomorphic to ?. However we can't witness this
-- isomorphism yet until we postulate the ? and ? paths.
--
-- unitPath' : (? ? ?) ? ?
-- unitPath' = ?

-- functions and types should have their equivalences as paths NOT
-- extensional isomorphisms which are later postulated to be paths!!!

-- need some kind of ``vertical'' path that allows us to move one level up.

-- explore this: 1 -> 1/x * x goes to point x which has a higher
-- groupoid structure ``above it'', gives you the higher groupoid
-- structure as 1/x and the naked point as x. Epsilon can then
-- recombine a higher groupoid collection of paths with a
-- point. Recall the 3/4 example: the paths not attached to any points
-- are 3/4: they take 4 points and return 3 points.

-- to get 1/5
-- start with 5 points * * * * *
-- and paths that connect them so that they are identified to be 1 point
-- Separate the points from the paths: the paths represent 1/5: something that takes
-- 5 points and collapses them to 1 and so we get 1_{5} --> 1/5 * 5
-- The problem is that 5 now has no paths over it so it's a different 5 and 1/5 has
-- no points so it's not a groupoid or is it?

-- 1/x can switch the two components of [_,_] so that it is an involution

We don't have pointed types. We have HITs. Each type is a collection
of points and a collection of non-trivial paths. (We get refl, !, ?,
associativity, etc. for tree and we also get the higher groupoid
structure for free which witnesses for example the pentagon and
triangle axioms, i.e., the coherence of assoc and the nice interplay
with !).

Start with the type 1 at 3 [{a,b,c) , {(a==b),(b==c)}]. This type
represents ``1'' but uses three identical elements to express this fact.

The paths by themselves are some sort of function that takes three
points and identifies them; this can be represented by the type
1/3. The three points by themselves can be represented by the type 3.

So 3 is [{a,b,c}, ?]

and 1/3 is [{? a b c ? (a==b,b==c)}, ?] 
the ? is a mini one step function to express that the paths are not
attached to any end points. You give me the end points and I will connect them.

The only value of type 1/3 is that function. We can pass around it and
eventually we must apply to 3 points which become identified. So
eventually we recover 1 back.

So now what is 1/(1/3). Remember that to express 1/3. We started with
a version of 1 that had 3 elements and paths between them. So to do
1/(1/3) we need to start with a version of 1 that has 1/3 elements and
paths between them.

1 at 1/3 is [{? a b c ? (a==b,b==c)}, 3 paths from this point to itself ]

so 1 at 1/3 can be split into 1/3 and and 1/(1/3) as follows:
1/3 = [{? a b c ? (a==b,b==c)}, ?]
1/(1/3) = [ ? a ? 3 paths from a to itself , ? ]

--

swapP : {A B : Set} ? A × B ? B × A
swapP {A} {B} = ?

It looks like I will need an axiom (not univalence necessarily but
something like it. We want to say that type isos from Pi correspond to
paths. So instead of saying (A ? B) ? (A ? B) we would say (A ? B) ?
(A ? B) where ? is the type of isos from Pi (making the types into a
field like structure).

Ok so there is an external map between A × B and B × A but because we
are monoidal closed there is also an internal map (1/(A × B)) × (B × A)
or more readably (A × B) -o (B × A). This map is defined as follows.

First let's start with 1 at A × B. This is:

[ {(a0,b0),(a0,b1),(a1,b0),(a1,b1)} , {(a0,b0)?(a0,b1), (a0,b0)?(a1,b0), ... }]

Anyway the point is that we can build 1/(A×B). Now we know that every
(a0,b0)?(a1,b1) comes from (a0?a1) and (b0?b1) and hence we can
assemble them in the other direction to get (b0,a0)?(b1,a1).

Now we can split using ? to 1/(A×B) and (A×B) where:

A×B = [ {(a0,b0),(a0,b1),(a1,b0),(a1,b1)} , ?]
and
1/(A×B) = [? (a0,b0) (a0,b1) (a1,b0) (a1,b1) ? 
             {(a0,b0)?(a0,b1), (a0,b0)?(a1,b0), ... } , 
           ?]

which is the same as:

1/(A×B) = [? a0 a1 b0 b1 ? {a0?a1, b0?b1} , ?]

Now the map (A×B)-o(B×A) is represented by the type:

[{((b0,a0),? a0 a1 b0 b1 ? {a0?a1, b0?b1}),
   ((b0,a1),? a0 a1 b0 b1 ? {a0?a1, b0?b1}),
   ((b1,a0),? a0 a1 b0 b1 ? {a0?a1, b0?b1}),
   ((b1,a1),? a0 a1 b0 b1 ? {a0?a1, b0?b1})}
  ?]

Now applying the function to (a0,b0) means pairing the above with (a0,b0):

[{((b0,a0),(a0,b0),? a0 a1 b0 b1 ? {a0?a1, b0?b1}),
   ((b0,a1),(a0,b0),? a0 a1 b0 b1 ? {a0?a1, b0?b1}),
   ((b1,a0),(a0,b0),? a0 a1 b0 b1 ? {a0?a1, b0?b1}),
   ((b1,a1),(a0,b0),? a0 a1 b0 b1 ? {a0?a1, b0?b1})}
  ?]

--}

------------------------------------------------------------------------------
